Illusion Game Validation (Study 1)

### 2022-08-05 study0/analysis.Rmd
update stim

df <- filter(df, RT < 1000)
```

@@

```{r message=FALSE, warning=FALSE}
visualize <- function(data, model_err, model_rt) {
  model <- mgcv::gam(Error ~ s(Illusion_Difference, by = Illusion_Side),
                     data = data,
                     family = "binomial")

  p1 <- data |> 
    group_by(Illusion_Difference, Illusion_Side) |> 
    summarize(Error = mean(Error)) |> 
    ggplot(aes(x = Illusion_Difference, y = Error)) +
    geom_bar(stat="identity") +
    geom_line(data=estimate_relation(model), aes(y=Predicted), color="blue") +
    geom_line(data=estimate_relation(model_err), aes(y=Predicted), color="red") +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    theme_modern() +
    labs(
      color = "Illusion Strength", 
      fill = "Illusion Strength",
      y = "Probability of Error",
      x = "Task Difficulty"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5)) + 
    facet_wrap(~Illusion_Side, ncol=2, labeller = "label_both")

  model <- mgcv::gam(RT ~ s(Illusion_Difference, by = Illusion_Side),
                     data = filter(data, Error == 0, RT < 1000))

  p2 <- data |> 
    filter(RT < 1000) |> 
    ggplot(aes(x = Illusion_Difference, y = RT)) +
    geom_point() +
    geom_line(data=estimate_relation(model), aes(y=Predicted), color="blue") +
    geom_line(data=estimate_relation(model_rt), aes(y=Predicted), color="red") +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme_modern() +
    labs(
      color = "Illusion Strength", 
      fill = "Illusion Strength",
      y = "RT",
      x = "Task Difficulty"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5)) + 
    facet_wrap(~Illusion_Side, ncol=2, labeller = "label_both")

@@

model_err <- glmmTMB::glmmTMB(Error ~ log(Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(Illusion_Difference) + (1|Participant),  
                              data = data)

@@

>>>>

df <- filter(df, RT < 900)
```

@@

```{r message=FALSE, warning=FALSE}
visualize <- function(data, model_err, model_rt) {

  at <- list(Illusion_Difference=seq(0, max(data$Illusion_Difference), length.out=30),
             Illusion_Side = levels(unique(data$Illusion_Side)))

  model <- mgcv::gamm(Error ~ s(Illusion_Difference, by = Illusion_Side),
                      random = list(Participant = ~1) ,
                      data = data,
                      family = "binomial")
  model_lin <- glmmTMB::glmmTMB(Error ~ Illusion_Difference + (1|Participant) ,
                      data = data,
                      family = "binomial")

  p1 <- data |> 
    group_by(Illusion_Difference, Illusion_Side) |> 
    summarize(Error = mean(Error)) |> 
    ggplot(aes(x = Illusion_Difference, y = Error)) +
    geom_bar(stat="identity") +
    geom_hline(yintercept=0.5, linetype="dotted") +
    geom_line(data=estimate_relation(model_lin, at=at[1]), aes(y=Predicted), color="black") +
    geom_line(data=estimate_relation(model, at=at), aes(y=Predicted), color="blue") +
    geom_line(data=estimate_relation(model_err, at=at[1]), aes(y=Predicted), color="red") +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    theme_modern() +
    labs(
      color = "Illusion Strength", 
      fill = "Illusion Strength",
      y = "Probability of Error",
      x = "Task Difficulty"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5)) + 
    facet_wrap(~Illusion_Side, ncol=2, labeller = "label_both")

  model <- mgcv::gamm(RT ~ s(Illusion_Difference, by = Illusion_Side),
                     random = list(Participant = ~1) ,
                     data = filter(data, Error == 0, RT < 1000))
    model_lin <- glmmTMB::glmmTMB(RT ~ Illusion_Difference + (1|Participant) ,
                      data = data)

  p2 <- data |> 
    filter(RT < 1000) |> 
    # group_by(Illusion_Difference, Illusion_Side) |> 
    # summarize(RT = mean(RT)) |> 
    ggplot(aes(x = Illusion_Difference, y = RT)) +
    ggdist::stat_pointinterval(aes(group=Illusion_Difference)) +
    geom_line(data=estimate_relation(model_lin, at=at[1]), aes(y=Predicted), color="black") +
    geom_line(data=estimate_relation(model, at=at), aes(y=Predicted), color="blue") +
    geom_line(data=estimate_relation(model_rt, at=at[1]), aes(y=Predicted), color="red") +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme_modern() +
    labs(
      color = "Illusion Strength", 
      fill = "Illusion Strength",
      y = "RT",
      x = "Task Difficulty"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5)) + 
    facet_wrap(~Illusion_Side, ncol=2, labeller = "label_both")

@@

model_err <- glmmTMB::glmmTMB(Error ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data)

@@

```{r message=FALSE, warning=FALSE}
optimal <- function(model_err) {
  pred <- estimate_relation(model_err, length=500)

  bounds <- pred |> 
      slice(c(which.min(abs(Predicted - 0.25)), 
              which.min(abs(Predicted - 0.025))))

  rez <- bounds |> 
    transmute(Illusion_Difference_log = Illusion_Difference_log,
              Illusion_Difference = exp(Illusion_Difference_log) - 1,
              Error = insight::format_value(Predicted, as_percent=TRUE))

  p <- pred |> 
    ggplot(aes(y = Predicted, x = Illusion_Difference_log)) +
    geom_line() +
    geom_hline(yintercept = c(0.25, 0.025), linetype="dotted", color=c("red", "green")) +
    geom_vline(xintercept = bounds$Illusion_Difference_log, linetype="dotted", color=c("red", "green")) +
    theme_modern()
  list(rez=rez, p=p)
}

model_err <- glmmTMB::glmmTMB(Error ~ Illusion_Difference_log + (1|Participant),  
                              data = data, family = "binomial")

rez <- optimal(model_err)
rez$rez
rez$p
```

### Müller-Lyer

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "Müller-Lyer")
```

```{r message=FALSE, warning=FALSE}
best_models(data)
```

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data)

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ Illusion_Difference_log + (1|Participant),  
                              data = data, family = "binomial")

rez <- optimal(model_err)
rez$rez
rez$p
```

### Vertical-Horizontal

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "Vertical-Horizontal")
```

```{r message=FALSE, warning=FALSE}
best_models(data)
```

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ sqrt(Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ sqrt(Illusion_Difference) + (1|Participant),  
                              data = data)

### Poggendorff

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "Poggendorff")
```

```{r message=FALSE, warning=FALSE}
best_models(data)
```

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ log(Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(Illusion_Difference) + (1|Participant),  
                              data = data)

### Zöllner

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "Zöllner")
```

```{r message=FALSE, warning=FALSE}
best_models(data)
```

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ log(Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(Illusion_Difference) + (1|Participant),  
                              data = data)

### Rod-Frame

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "Rod-Frame")
```

```{r message=FALSE, warning=FALSE}
best_models(data)
```

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ log(Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(Illusion_Difference) + (1|Participant),  
                              data = data)

### Contrast

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "Contrast")
```

```{r message=FALSE, warning=FALSE}
best_models(data)
```

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data)

### White

```{r message=FALSE, warning=FALSE}
data <- dplyr::filter(df, Illusion_Type == "White")
```

```{r message=FALSE, warning=FALSE}
best_models(data)
```

```{r message=FALSE, warning=FALSE}
model_err <- glmmTMB::glmmTMB(Error ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data, family = "binomial")
model_rt <- glmmTMB::glmmTMB(RT ~ log(1+Illusion_Difference) + (1|Participant),  
                              data = data)

### 2022-08-05 study1.Rmd
update stim

```{r message=FALSE, warning=FALSE, cache=cache}
# plot(seq(0, 1, length.out=100), seq(0, 1, length.out=100)**(1/3), "l", col="green")
# lines(seq(0, 1, length.out=100), sqrt(seq(0, 1, length.out=100)), "l", col="red")
# lines(seq(0, 1, length.out=100), log(1+seq(0, 1, length.out=100)), "l", col="blue")
# lines(seq(0, 1, length.out=100), seq(0, 1, length.out=100), "l", col="black")

best_models <- function(data) {
  models <- list()
  for(i in 1:1) {
    for(j in 1:1) {
      for(k1 in c("", "_log", "_sqrt", "_cbrt")) { 
        for(k2 in c("", "_log", "_sqrt", "_cbrt")) {  
          for(side in c("", "-side")) {
            name <- paste0("dif", k1, i, "-", "str", k2, j, side)
            print(name)
            f <- paste0("poly(Illusion_Difference", 
                        k1,
                        ", ",
                        i,
                        ") * poly(Illusion_Strength",
                        k2, 
                        ", ",
                        j, 
                        ") + (1|Participant)")

@@

```{r message=FALSE, warning=FALSE, cache=cache}
cbrt <- function(x) sign(x) * abs(x)^(1/3)

formula <- brms::bf(
  Error ~ poly(Illusion_Difference, 2) * Illusion_Strength + 
    (1 | Participant),
  family = "bernoulli"
)

@@

  n_lines <- length(unique(data[[vardiff]]))
  pred <- estimate_relation(model,
                            at = vars,
                            length = c(NA, 25))
  pred$Illusion_Difference <- as.factor(round(pred[[vardiff]], 2))

  # Set colors for lines
  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(n_lines)
  diffvals <- as.numeric(as.character(unique(sort(pred$Illusion_Difference))))
  names(colors) <- diffvals

@@

  # Manual jittering
  xrange <- 0.05*diff(range(data$Illusion_Strength))
  data$x <- data$Illusion_Strength 
  data$x[data$x > 0] <- data$x[data$x > 0] - runif(sum(data$x > 0), 0, xrange)
  data$x[data$x < 0] <- data$x[data$x < 0] + runif(sum(data$x < 0), 0, xrange)
  data$x[round(data$x, 2) == 0] <- data$x[round(data$x, 2) == 0] + runif(sum(round(data$x, 2) == 0), -xrange/2, xrange/2)

  pred |>
    ggplot(aes(x = Illusion_Strength, y = Predicted)) +
    geom_dots(
      data = mutate(data, 
                    Illusion_Strength = round(Illusion_Strength, 1)),
      aes(x=x, y = Error, group = Error, side = .dots_side, order=color), 
      fill = data$color,
      color = NA, 
      alpha=0.5) +
    geom_slab(data=data, aes(y = Error)) +
    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Illusion_Difference, group = Illusion_Difference), alpha = 0.2) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_hline(yintercept = c(0.05, 0.5, 0.95), linetype = "dotted", alpha = 0.5) +
    geom_line(aes(color = Illusion_Difference, group = Illusion_Difference)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_color_manual(values = colors) +
    scale_fill_manual(values = colors) +
    coord_cartesian(xlim=c(min(data$Illusion_Strength), max(data$Illusion_Strength))) +
    theme_modern() +
    labs(
      title = paste0(data$Illusion_Type[1], " Illusion"),
      color = "Difficulty", fill = "Difficulty",
      y = "Probability of Error",
      x = "Illusion Strength"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))
}

@@

@@

parameters::parameters(model)
```

@@

>>>>

```{r message=FALSE, warning=FALSE, cache=cache}
best_models <- function(data) {
  models <- list()
  for(i in 1:1) {
    for(j in 1:1) {
      for(k1 in c("", "_log", "_sqrt", "_cbrt")) { 
        for(k2 in c("")) {  
          for(side in c("", "-side")) {
            name <- paste0("dif", k1, i, "-", "str", k2, j, side)
            # print(name)
            f <- paste0("poly(Illusion_Difference", 
                        k1,
                        ", ",
                        i,
                        ") * poly(Illusion_Strength",
                        k2, 
                        ", ",
                        j, 
                        ") + (1|Participant)")

@@

```{r message=FALSE, warning=FALSE, cache=cache}
cbrt <- function(x) sign(x) * abs(x)**(1/3)

formula <- brms::bf(
  Error ~ cbrt(Illusion_Difference) * Illusion_Strength + 
    (1 | Participant),
  family = "bernoulli"
)

@@

  # Get predicted
  pred <- estimate_relation(model,
                            at = vars,
                            length = c(NA, 25))
  pred[[vardiff]] <- as.factor(pred[[vardiff]])

  # Set colors for lines
  colors <- colorRampPalette(c("#F44336", "#FFC107", "#4CAF50"))(length(unique(data[[vardiff]])))
  diffvals <- as.numeric(as.character(unique(sort(pred[[vardiff]]))))
  names(colors) <- diffvals

@@

  # Manual jittering
  xrange <- 0.05*diff(range(data[[varstrength]]))
  data$x <- data[[varstrength]]
  data$x[data$x > 0] <- data$x[data$x > 0] - runif(sum(data$x > 0), 0, xrange)
  data$x[data$x < 0] <- data$x[data$x < 0] + runif(sum(data$x < 0), 0, xrange)
  data$x[round(data$x, 2) == 0] <- data$x[round(data$x, 2) == 0] + runif(sum(round(data$x, 2) == 0), -xrange/2, xrange/2)

  pred |>
    ggplot(aes_string(x = varstrength, y = "Predicted")) +
    geom_dots(
      data = data,
      aes(x=x, y = Error, group = Error, side = .dots_side, order=color), 
      fill = data$color,
      color = NA, 
      alpha=0.5) +
    geom_slab(data=data, aes(y = Error)) +
    geom_ribbon(aes_string(ymin = "CI_low", ymax = "CI_high", fill = vardiff, group = vardiff), alpha = 0.2) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_hline(yintercept = c(0.05, 0.5, 0.95), linetype = "dotted", alpha = 0.5) +
    geom_line(aes_string(color = vardiff, group = vardiff)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0), labels = scales::percent) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_color_manual(values = colors) +
    scale_fill_manual(values = colors) +
    coord_cartesian(xlim=c(min(data[[varstrength]]), max(data[[varstrength]]))) +
    theme_modern() +
    labs(
      title = paste0(data$Illusion_Type[1], " Illusion"),
      color = "Difficulty", fill = "Difficulty",
      y = "Probability of Error",
      x = "Illusion Strength"
    ) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))
}

@@

```{r message=FALSE, warning=FALSE, cache=cache}
formula <- brms::bf(
  Error ~ cbrt(Illusion_Difference) * Illusion_Strength + 
    (1 | Participant),
  family = "bernoulli"
)

@@

@@

# parameters::parameters(model)
```
